---
title: "p8105_hw1_xc2610"
author: "Xuanhe Chen"
date: 2019-09-21
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

Problem 1

Create the dataframe accordingly:
```{r df_create}
set.seed(5)

random_df = tibble(
  val = rnorm(10, sd = .5), ## random sample of size 10
  vec_logical = c(TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE), ## initialize the logical factor as all true
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"), ## character vector of length 10
  vec_factor = factor(c("pro", "con", "non", "non", "con", "pro", "con", "non", "pro", "non")) ## factor vector of length 10 with 3 different levels
)

## show(random_df)

random_df$vec_logical <- ifelse(random_df$val < 0, FALSE, TRUE) ## edit the logical vector according to positive or not of each sample value

## show(random_df)
```

Then try to get the mean of each variables, mean only works on the sample value and logical factor:
```{r mean}
mean(pull(random_df, val)) ## try get mean of values
mean(pull(random_df, vec_logical)) ## try get mean of logical factors
mean(pull(random_df, vec_char)) ## try get mean of characters
mean(pull(random_df, vec_factor)) ## try get mean of factors
```
 as.numertic works on vec_logical and vec_factor but not vec_char. Booleans are considered as 1s and 0s so the mean of a list of booleans can be calculated. Characters can not be shown as numeric values so we can't calculate its mean. We can't get the mean of factors although it can be shown as numertic values might because each facotr name doesn't map to some particular (like 0 is always False 1 is always true). 

```{r as.numeric}
## convert variables to numeric
as.numeric(pull(random_df, vec_logical))
as.numeric(pull(random_df, vec_char))
as.numeric(pull(random_df, vec_factor))
``` 
Problem 2

  This dataset contains 8 variables: "species", "island", "bill_length_mm", bill_depth_mm", "flipper_length_mm", "body_mass_g", "sex" and "year". "species", "island", "gender" and "year" are categorical, while "bill_length_mm", bill_depth_mm", "flipper_length_mm" and "body_mass_g" are numerical. The data frame contains 344 rols and 8 columns. 
  
  The mean flipper lengths of penguins according to this data is 199.7471mm.
```{r penguins}
## get the penguins dataset
data("penguins", package = "palmerpenguins")

## check information about this dataset
names(penguins)
## view(penguins)
head(penguins)
nrow(penguins)
ncol(penguins)
## skimr::skim(penguins)
```

```{r mean_flipper_length}
penguins$flipper_length_mm [is.na(penguins$flipper_length_mm)] <- 0 ## rewite all NA's as 0
## head(penguins)
mean(pull(penguins, flipper_length_mm)) ## calculate the mean length
```

Make a scatterplot of flipperlength vs billlength color points and assign different color to each species:
```{r scatter}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() ## create a scatter plot according to penguin species
## save the scatter plot to local
## ggsave("scatter_HW1_P2.pdf", height = 4, width = 6)
```